---
title: "quiz5"
author: "Shengnan You"
date: "4/7/2020"
output: html_document
---
#10

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data}
setwd("~/Desktop")
library('MASS')
data <- read.table("banknote.txt",header=FALSE,sep=',')
head(data)
colnames(data) <- c('variance','skewness','curtosis','entropy','class')
```
```{r Q1}
full <- glm(class ~ ., family = 'binomial', data = data)
summary(full)
step_model1 <- stepAIC(full,trace = TRUE,k = log(nrow(data)))
summary(step_model1)
BIC(step_model1)
```
Our final model is class ~ variance + skewness + curtosis. BIC :82.19474


```{r Q2}
full2 <- glm(class ~ .^2, family = 'binomial' , control = list(maxit =50), data = data)
summary(full2)
suppressWarnings(
step_model2 <- stepAIC(full2, scope = list(lower=~1,upper=full2),k = log(nrow(data)), trace = FALSE)
)
summary(step_model2)
BIC(glm(class ~ variance + skewness + curtosis + variance:curtosis, 
    family = "binomial", data = data))

```

The final model is class ~ variance + skewness + curtosis + variance:curtosis and the BIC is 36.12012.

```{r Q3}

library(leaps)
library(bestglm)
library(dummies)
best_model3 <- bestglm(data, family = binomial, IC = 'BIC')
best_model3
BIC(glm(class ~ variance + skewness + curtosis , data = data, family = 'binomial'))
```
Our model is class ~ variance + skewness + curtosis
BIC is 82.19474

```{r Q4}
interactions <-data.frame(model.matrix(class ~ (variance + skewness + curtosis +entropy)^2,data = data))
head(interactions)
interactions <- interactions[,-1]
interactions$class <- data$class
suppressWarnings(
best_model4 <- bestglm(interactions, family = binomial, IC = 'BIC')
)
summary(best_model4)
BIC(glm(class ~ variance + skewness + curtosis + variance:curtosis, 
    family = "binomial", data = data))
```
Our final model is class ~ variance + skewness + curtosis + variance.curtosis
BIC is 36.12012

Q5
The best model is class ~ variance + skewness + curtosis + variance:curtosis which has the lowest BIC




